---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(here)
library(tidyboot)
library(lme4)
library(lmerTest)
library(ggthemes)
```

```{r}
d.targetData <- read_csv(here('data/comprehension/post_test_utterances.csv'))
d.comprehension.raw <- read_csv(here('data/comprehension/dataFromMongo.csv')) %>%
  filter(iterationName != 'full_sample')
  
passed_catch <- d.comprehension.raw %>%
  filter(tangram == 'catch' & trial_type == 'html-button-response') %>%
  filter(response == 'G') %>%
  pull(wID)

complete_data <- d.comprehension.raw %>%
  group_by(wID) %>%
  summarize(final_trial = last(trial_index)) %>%
  filter(final_trial >= 21) %>%
  pull(wID)

d.comprehension <- d.comprehension.raw %>%
  filter(wID %in% passed_catch) %>%
  filter(wID %in% complete_data) %>%
  filter(!is.na(tangram)) %>%
  ungroup() %>%
  mutate(correct = response == tangram,
         logrt = log(rt),
         upper = mean(logrt) + 2*sd(logrt),
         lower = mean(logrt) - 2*sd(logrt)) %>%
  filter(logrt < upper, logrt > lower)
```

Look at how many annotations we got for each description.

```{r}
d.targetData %>%
  mutate(tangram = substr(tangram, 21, 21)) %>%
  rename(targetAudience = ownGroup) %>%
  left_join(d.comprehension %>%
    group_by(playerId, tangram, origText) %>%
    filter(wID %in% passed_catch) %>%
    filter(wID %in% complete_data) %>%
    tally()) %>%
  group_by(n) %>%
  tally() %>%
  ggplot(aes(x = n/2, y = nn)) +
    geom_bar(stat = 'identity') +
    theme_few() +
    labs('# tangrams with annotations', '# annotations')
```

Look at response time distribution to get a sense of outliers.

```{r}
d.comprehension %>%
  ggplot(aes(x = logrt)) +
    geom_histogram() +
    geom_vline(aes(xintercept = lower)) +
    geom_vline(aes(xintercept = upper))
```
Implement exclusions and pre-process data frame

```{r}
d.4afc <- d.comprehension %>%
  filter(trial_type == 'html-button-response') %>%
  filter(tangram != 'catch') %>%
  left_join(d.targetData, by = c('playerId', 'tangram', 'origText', 'gameId')) %>%
  select(-targetAudience) %>%
  rename(targetAudience = ownGroup) %>%
  mutate(condition = case_when(
    condition == 'ownGroup_second' ~ 'otherGroup_first', 
    condition == 'otherGroup_second' ~ 'ownGroup_first',
    TRUE ~ condition)
  ) %>%
  filter(!is.na(condition))
```

```{r}
dodge <- position_dodge(width = 0.2)
d.4afc %>%
  select(playerId, tangram, condition, targetAudience, correct, logrt) %>%
  mutate(correct = 100 * correct) %>%
  gather(metric, value, correct, logrt) %>%
  group_by(metric, condition, targetAudience) %>%
  tidyboot_mean(value, na.rm = T) %>%
  ggplot(aes(x = condition, y = empirical_stat, color = targetAudience)) +
    geom_point(position= dodge) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                  width = 0, position= dodge) +
    theme_few() +
    # scale_x_discrete(labels=c("first"="other -> own",
    #                           "second"="own -> other")) +
    labs(x = '', y = '% correct') +
    facet_wrap(~ metric, scales = 'free') +
    theme(aspect.ratio = 1, 
          axis.text.x = element_text(angle = 45,  vjust = 1, hjust=1))

ggsave('../writing/exp2_results_raw.pdf', units = 'in', height = 3, width = 4)
```

Not controlling for clustered variance, it looks like we just have a (very) weak simple effect of audience for participants where the 'other group' question was asked first.

```{r}
d.4afc %>%
  glm(correct ~ targetAudience * condition,
       family = 'binomial',
       data = .) %>%
  summary()
```

Controlling for clustered variance at the annotator level and the source speaker level, the simple effect of audience remains significant but we also get a marginal interaction.

```{r}
d.4afc %>%
  select(gameId, wID, playerId, tangram, origText, targetAudience, condition, correct, blockOrder) %>%
  group_by(gameId, playerId, tangram, targetAudience, condition) %>%
  summarize(prop = mean(correct), n = length(correct)) %>%
  glmer(prop ~ targetAudience * condition  
        + (1 + condition | tangram)
        + (1 + condition | gameId)
        + (1 | playerId),
        weights = n,
       family = 'binomial',
       data = .) %>%
  summary()
```

```{r}
d.4afc %>%
  brms::brm(correct ~ targetAudience * condition  
                + (1 | wID)
                + (1 + targetAudience * condition | tangram) 
                + (1 + targetAudience | playerId),
       family = 'bernoulli',
       data = .) %>%
  summary()
```


```{r}
d.4afc %>%
  lmer(logrt ~ targetAudience * condition  
             + (1 + targetAudience | playerId) 
             + (1 | tangram)
             + (1 + targetAudience + condition | wID)
             + (1 | gameId),
        control = lmerControl(optimizer = 'bobyqa'),
       data = .) %>%
  summary()
```


```{r}
d.slider <- d.comprehension %>%
  filter(wID %in% passed_catch) %>%
  filter(!is.na(tangram)) %>%
  filter(tangram != 'catch') %>%
  select(wID, rt, origText, response, tangram, gameId, playerId, trial_type) %>%
  left_join(d.targetData %>% 
              mutate(tangram = substr(tangram, 21, 21)) %>% 
              rename(targetAudience = ownGroup),
            by = c('playerId', 'tangram', 'origText', 'gameId')) %>%
  mutate(condition = case_when(
    condition == 'ownGroup_second' ~ 'otherGroup_first', 
    condition == 'otherGroup_second' ~ 'ownGroup_first',
    TRUE ~ condition)
  ) %>%
  pivot_wider(names_from = trial_type, values_from = c('response', 'rt')) %>%
  unnest() %>%
  mutate(correct = `response_html-button-response` == tangram,
         `response_html-slider-response` = as.numeric(`response_html-slider-response`))

d.slider %>%
  group_by(correct, targetAudience, condition) %>%
  tidyboot_mean(`response_html-slider-response`) %>%
  ggplot(aes(x = condition, y = empirical_stat, color = targetAudience)) +
    geom_point(position = dodge) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
    facet_wrap(~ correct) +
    theme_few()
```

```{r}
d.cor <- d.4afc %>% 
  group_by(gameId, playerId, tangram, targetAudience) %>% 
  summarize(correct = mean(correct), rt = mean(rt), uttLength = mean(uttLengthWords)) %>% 
  pivot_wider(names_from = targetAudience, values_from = c('correct', 'rt', 'uttLength')) %>% 
  mutate(correct_diff = correct_otherGroup -correct_ownGroup,
         rt_diff = rt_otherGroup - rt_ownGroup,
         uttLength_diff = uttLength_otherGroup - uttLength_ownGroup) %>%
  select(gameId, playerId, tangram, ends_with('diff'))
  
cor.test(d.cor$uttLength_diff, d.cor$correct_diff, use="complete.obs")
```

Plot relationship... 

```{r}
d.cor %>%
  ggplot(aes(x = scale(uttLength_diff), y = scale(correct_diff))) +
    geom_jitter(width = .5, height = .5) +
    geom_smooth(method = 'lm')
```
