---
title: "Experiment 1 analysis"
output: html_notebook
---


```{r}
library(tidyverse)
library(jsonlite)
library(here)
library(lme4)
library(lmerTest)
library(ggthemes)
library(tidyboot)

source('helpers.R')
```

```{r}
d.games <- read_csv(here('data/games.csv')) %>% 
  rename(gameId = `_id`)

d.chat <- read_csv(here('data/rounds.csv')) %>%
  filter(createdAt >= lubridate::ymd('2021-01-21')) %>%
  mutate(data.chat = ifelse(is.na(data.chat), '{}', data.chat)) %>%
  rename(row_id = `_id`) %>%
  mutate(data.chat = map(data.chat, .f = ParseJSONColumn)) %>%
  unnest(data.chat) %>%
  select(-data.target, -ends_with('response'), -ends_with('correct')) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE))

d.room_info <- read_csv(here('data/rounds.csv')) %>%
  filter(createdAt >= lubridate::ymd('2021-01-21')) %>%
  mutate(data.target = map(data.target, .f = ParseJSONColumn)) %>%
  unnest(data.target) %>%
  rename(room0_target = room0, room1_target = room1) %>%
  filter(!is.na(text)) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
  rename(room1_response = room1response, room0_response = room0response) %>%
  select(-room0data, -room1data, -ends_with('correct'), -chat) %>%
  gather(key, value, starts_with('room')) %>%
  separate(key, into = c('roomId', 'info')) %>%
  spread(info, value) %>%
  mutate(response = ifelse(response == 'false', 'timed_out', response))
```

Get exit survey responses

```{r}
d.exit.survey <- read_csv(here('data/player-inputs.csv')) %>%
  filter(createdAt >= lubridate::ymd('2021-01-21')) %>%
  filter(!is.na(data.community)) %>%
  left_join(d.games, by = c('gameId')) %>%
  select(where(~ !any(is.na(.))))

correct_community_response <- d.exit.survey %>%
  filter(data.community == data.teamColor) %>%
  pull(playerId)
```

```{r}
roles.tmp <- d.games %>%
  select(gameId, data.roleList) %>%
  mutate(data.roleList = map(data.roleList, .f = ParseJSONColumn)) %>%
  unnest(data.roleList) %>%
  gather(playerId, role, -gameId) %>%
   rowwise() %>%
  filter(!is.null(role)) %>%
  unnest(role) %>%
  group_by(playerId) %>%
  mutate(n = row_number() - 1,
         trialNum = n %% 16,
         partnerNum = floor(n / 16),
         repNum = floor(trialNum/4) %% 4) %>%
  select(-n)

roles <- d.chat %>%
  ungroup() %>%
  mutate(row_id = row_number()) %>%
  group_by(gameId, partnerNum, roomId, trialNum, repNum) %>%
  filter(row_id == first(row_id)) %>%
  group_by( gameId, partnerNum, roomId) %>%
  summarize(playerId = unique(playerId)) %>%
  left_join(roles.tmp) %>%
  group_by(gameId, partnerNum, roomId, trialNum, repNum) %>%
  filter(!is.na(roomId)) %>%
  spread(role, playerId)
```

Some games had a number of rounds go by where no one talked (one possible upstream reason for timeouts...)

```{r}
talked <- d.chat %>%
  filter(!is.na(text)) %>%
  group_by(gameId, trialNum, roomId, partnerNum) %>%
  tally() %>%
  group_by(gameId) %>%
  tally() %>%
  filter(n >= 90) %>%
  pull(gameId)
```

```{r}
post_test.raw <- read_csv(here('data/player-inputs.csv')) %>%
  filter(gameId %in% talked) %>%
  filter(playerId %in% correct_community_response) %>%
  group_by(playerId) %>%
  mutate(blockOrder = ifelse(row_number() <= 4, 'first', 'second')) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
  mutate(origText = gsub("\\n", '', fixed = T, text),
         text = gsub("[/?/.]", ' ', origText),
         text = str_squish(text),
         utt_length_chars = str_length(text), 
         utt_length_words = str_count(text, "\\W+") + 1,
         ownGroup = ifelse(ownGroup == group, 'ownGroup', 'otherGroup')) %>%
  filter(!is.na(text))

full_post_test <- post_test.raw %>%
  group_by(gameId) %>% 
  tally() %>%
  filter(n == 32) %>%
  pull(gameId)

cat('we have', length(full_post_test), 'networks where all 4 participants gave post-test responses.\n')
cat("that's", length(unique(post_test.raw$playerId)), 'individual participants')
```

# Behavioral results during game 

accuracy over time.

```{r}
d.room_info %>%
  filter(gameId %in% talked) %>%
  select(gameId, trialNum, partnerNum, repNum, roomId, response, target) %>%
  # Remove trials that timed out (so not confounded by attention effects)
  filter(response != 'timed_out') %>%
  mutate(correct = response == target) %>%
  group_by(partnerNum, repNum) %>%
  tidyboot::tidyboot_mean(correct) %>%
  ggplot(aes(x = repNum, y = empirical_stat)) +
    geom_line() +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0) +
    facet_grid(. ~ partnerNum) +
    theme_few()
```

Utterance length over time.

```{r}
dodge <- position_dodge(width = 0.5)

final_block <- post_test.raw %>%
  group_by(blockOrder, ownGroup) %>%
  tidyboot_mean(utt_length_words) %>%
  mutate(partnerNum = 3,
         repNum = ifelse(blockOrder == 'first', 0, 1)) %>%
  ungroup() %>%
  select(-blockOrder)

d.chat %>%
  filter(gameId %in% talked) %>%
  filter(role == 'speaker') %>%
  mutate(text = gsub("\\n", '', fixed = T, text),
         text = gsub("[/?/.]", ' ', text),
         text = str_squish(text),
         utt_length_chars = str_length(text), 
         utt_length_words = str_count(text, "\\W+") + 1) %>%
  group_by(gameId, trialNum, partnerNum, repNum, roomId) %>%
  summarize(total_num_words = sum(utt_length_words),
            total_num_chars = sum(utt_length_chars)) %>%
  group_by(partnerNum, repNum) %>%
  tidyboot_mean(total_num_words) %>%
  bind_rows(final_block) %>%
  ggplot(aes(x = repNum + 1, y = empirical_stat, group = ownGroup, color = ownGroup)) +
    geom_line(position = dodge) +
    geom_point(position = dodge) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0, position = dodge) +
    facet_grid(. ~ partnerNum) +
    theme_few() +
    theme(aspect.ratio = 2)

ggsave('../writing/speaker_curves.pdf', units = 'in', height = 3, width = 6)
```

```{r}
post_test <- post_test.raw %>%
  filter(gameId %in% full_post_test) %>%
  select(gameId, playerId, tangram, utt_length_words, ownGroup, blockOrder) %>%
  spread(ownGroup, utt_length_words) %>%
  mutate(diff = log(otherGroup / ownGroup)) %>%
  left_join(post_test.raw %>% 
      unite(condition, ownGroup, blockOrder) %>% 
      group_by(playerId, tangram) %>% 
      summarize(condition = first(condition)))
```

## description length longer for out-group

```{r}
post_test %>%
  gather(ownGroup, utt_length_words, otherGroup, ownGroup) %>%
  group_by(playerId) %>%
  # z score
  mutate(log_length = scale(log(utt_length_words))) %>%
  ggplot(aes(x = blockOrder, y = log_length, color = ownGroup)) +
    geom_point(position = position_jitter(width = .2, height = 0.05), 
               size = .1, alpha = 0.1) +
    geom_boxplot(aes(x = blockOrder, color = ownGroup, y = log_length),
                 outlier.shape = NA, alpha = 0.5, width = .6) +
    geom_hline(yintercept = 0, linetype = 'dotted') +
    theme_few() +
    scale_x_discrete(labels=c("first"="other -> own",
                              "second"="own -> other")) +
    labs(x = '', y = 'log # words (z scored)') +
    theme(aspect.ratio = 2, axis.text.x = element_text(angle = 45,  vjust = 1, hjust=1))

ggsave('../writing/utt_length.pdf', units = 'in', height = 3, width = 5)
```

maximal lmer model that converges (we get boundary warning but can ignore)

```{r}
regular.mod <- post_test.raw %>%
  filter(gameId %in% full_post_test) %>%
  lmer(log(utt_length_words) ~ ownGroup + blockOrder
                               + (1 + ownGroup + blockOrder | playerId)
                               + (1 | gameId)
                               + (1 | tangram),
       data = .)

summary(regular.mod)
```

```{r}
interaction.mod <- post_test.raw %>%
  filter(gameId %in% full_post_test) %>%
  lmer(log(utt_length_words) ~ ownGroup * blockOrder
                               + (1 + ownGroup * blockOrder | playerId)
                               + (1 | gameId)
                               + (1 | tangram),
       data = .)
anova(regular.mod, interaction.mod)
```

check boundary warning against full-blown brms model

```{r}
library(lme4)
library(lmerTest)
library(brms)
post_test.raw %>%
  filter(gameId %in% full_post_test) %>%
  select(gameId, playerId, tangram, utt_length_words, ownGroup, blockOrder) %>%
  brms::brm(log(utt_length_words) ~ ownGroup + blockOrder
                               + (1 + ownGroup + blockOrder | playerId)
                               + (1 + ownGroup + blockOrder | gameId)
                               + (1 + ownGroup + blockOrder | tangram),
       data = .) %>%
  summary()
```

```
Links: mu = identity; sigma = identity 
Formula: log(utt_length_words) ~ ownGroup + blockOrder + (1 + ownGroup + blockOrder | playerId) + (1 + ownGroup + blockOrder | gameId) + (1 + ownGroup + blockOrder | tangram) 
   Data: . (Number of observations: 1056) 
Samples: 4 chains, each with iter = 2000; warmup = 1000; thin = 1;
         total post-warmup samples = 4000

Group-Level Effects: 
~gameId (Number of levels: 33) 
                                       Estimate Est.Error l-95% CI
sd(Intercept)                              0.30      0.09     0.12
sd(ownGroupownGroup)                       0.06      0.05     0.00
sd(blockOrdersecond)                       0.05      0.04     0.00
cor(Intercept,ownGroupownGroup)            0.06      0.47    -0.82
cor(Intercept,blockOrdersecond)           -0.20      0.47    -0.91
cor(ownGroupownGroup,blockOrdersecond)    -0.05      0.50    -0.90
                                       u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)                              0.48 1.01      628      734
sd(ownGroupownGroup)                       0.17 1.00     1625     2048
sd(blockOrdersecond)                       0.15 1.00     1908     2694
cor(Intercept,ownGroupownGroup)            0.88 1.00     3295     2775
cor(Intercept,blockOrdersecond)            0.77 1.00     3991     3215
cor(ownGroupownGroup,blockOrdersecond)     0.87 1.00     3020     3024

~playerId (Number of levels: 132) 
                                       Estimate Est.Error l-95% CI
sd(Intercept)                              0.54      0.05     0.44
sd(ownGroupownGroup)                       0.16      0.06     0.04
sd(blockOrdersecond)                       0.15      0.06     0.03
cor(Intercept,ownGroupownGroup)            0.10      0.26    -0.37
cor(Intercept,blockOrdersecond)            0.02      0.28    -0.47
cor(ownGroupownGroup,blockOrdersecond)     0.64      0.28    -0.08
                                       u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)                              0.64 1.00     1505     2315
sd(ownGroupownGroup)                       0.27 1.00      998     1396
sd(blockOrdersecond)                       0.26 1.00      815     1356
cor(Intercept,ownGroupownGroup)            0.65 1.00     2352     2153
cor(Intercept,blockOrdersecond)            0.63 1.00     2522     2146
cor(ownGroupownGroup,blockOrdersecond)     0.97 1.00     1264     1498

~tangram (Number of levels: 8) 
                                       Estimate Est.Error l-95% CI
sd(Intercept)                              0.10      0.05     0.02
sd(ownGroupownGroup)                       0.06      0.05     0.00
sd(blockOrdersecond)                       0.04      0.04     0.00
cor(Intercept,ownGroupownGroup)            0.13      0.47    -0.77
cor(Intercept,blockOrdersecond)           -0.00      0.49    -0.86
cor(ownGroupownGroup,blockOrdersecond)     0.04      0.50    -0.86
                                       u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)                              0.22 1.00     1863     2034
sd(ownGroupownGroup)                       0.18 1.00     1512     1937
sd(blockOrdersecond)                       0.13 1.00     2858     2374
cor(Intercept,ownGroupownGroup)            0.89 1.00     4994     2805
cor(Intercept,blockOrdersecond)            0.87 1.00     6107     3120
cor(ownGroupownGroup,blockOrdersecond)     0.90 1.00     3834     3297

Population-Level Effects: 
                 Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS
Intercept            1.94      0.08     1.78     2.11 1.00     2550
ownGroupownGroup    -0.11      0.05    -0.20    -0.02 1.00     4135
blockOrdersecond    -0.13      0.04    -0.21    -0.05 1.00     4542
                 Tail_ESS
Intercept            2528
ownGroupownGroup     3015
blockOrdersecond     2997

Family Specific Parameters: 
      Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sigma     0.45      0.01     0.43     0.47 1.00     3290     2960
```

```{r}
post_test %>%
  ungroup() %>%
  summarize(result = list(t.test(.$ownGroup, .$otherGroup, paired = T))) %>%
  pull(result) 
```

```{r}
qualitative.direction <- post_test %>%
  group_by(gameId, playerId, ownGroup) %>%
  summarize(m = mean(utt_length_words)) %>%
  select(playerId, ownGroup, m) %>%
  spread(ownGroup, m) %>%
  mutate(qualitative = case_when(otherGroup > ownGroup ~ 'more',
                                 otherGroup == ownGroup ~ 'same',
                                 TRUE ~ 'less')) 

less_informative_to_other_group <- qualitative.direction %>%
  filter(qualitative=='less') %>%
  pull(playerId)

qualitative.direction %>%
  group_by(qualitative) %>%
  tally() %>%
  mutate(prop = n / sum(n))
```

# Supplemental

## Investigate timeouts... 

```{r}
time_outs <- d.room_info %>%
  left_join(roles) %>%
  group_by(listener, partnerNum, response) %>%
  tally() %>%
  filter(response == 'timed_out' | is.na(response)) %>%
  filter(!is.na(listener)) %>%
  group_by(listener, partnerNum) %>%
  summarize(totalTimeouts = sum(n))
```

Fewer timeouts as the game proceeds (need to be careful about this as a confound with accuracy).

```{r}
time_outs %>%
  ggplot(aes(x = totalTimeouts)) +
    geom_histogram(bins = 5) +
    geom_density(aes(y = ..count..), adjust = 3) +
    facet_wrap(~ partnerNum)
```

Overall distribution of total timeouts per participant.

```{r}
time_outs %>%
  group_by(listener) %>%
  summarize(totalTimeouts = sum(totalTimeouts))  %>%
  ggplot(aes(x = totalTimeouts)) +
    geom_histogram(bins = 24) +
    theme_few()
```

## Look at individual differences

```{r}
total_avg <- post_test %>%
  select(-diff) %>%
  gather(ownGroup, utt_length_words, otherGroup, ownGroup) %>%
  group_by(ownGroup) %>%
  tidyboot::tidyboot_mean(utt_length_words, nboot = 100) %>%
  select(-n, -mean) %>%
  pivot_wider(names_from = ownGroup, values_from = c('empirical_stat', 'ci_lower', 'ci_upper')) 

post_test %>%
  ggplot(aes(x = log(ownGroup), y = log(otherGroup), color = ownGroup > otherGroup)) +
    geom_point(position = position_jitter(width = 0.1, height = 0.1)) +
    geom_abline(aes(intercept = 0, slope = 1), linetype = 'dotted') +
    facet_wrap(~ condition) +
    ggthemes::theme_few() +
    ylim(0, 3.5) +
    xlim(0, 3.5) +
    labs(x = 'log # words for own group', y = 'log # words for other group') +
    theme(aspect.ratio = 1)
```

## Try to visualize participants as lines?

```{r}
error_bars <- post_test %>%
  group_by(ownGroup) %>%
  tidyboot::tidyboot_mean(utt_length_words, nboot = 100) %>%
  mutate(goingUp = TRUE)

post_test %>%
  mutate(goingUp = otherGroup > ownGroup) %>%
  gather(ownGroup, m, otherGroup, ownGroup) %>%
  ggplot(aes(x = ownGroup, y = log(m), color = goingUp)) +
    geom_line(aes(group = interaction(playerId, tangram)), alpha = 0.1) +
    geom_point(aes(x = ownGroup, y = log(empirical_stat)), 
               size = 5, data = error_bars) +
    geom_errorbar(aes(ymin = log(ci_lower), ymax = log(ci_upper), 
                      x = ownGroup, y = log(empirical_stat)), 
                  width = 0,
                  data = error_bars) +
    facet_wrap(~ condition) +
    geom_line(aes(x = ownGroup, group = 1, y = log(empirical_stat)), size= 2,data = error_bars) +
    theme_bw() +
    labs(x = '', y = 'log # words') +
    theme(aspect.ratio = 1)
```

# Vizualize the distribution of diffs in utterance length within participant

```{r}
summary.df <- post_test %>%
  group_by(playerId, condition) %>%
  summarize(diff = mean(diff)) %>%
  group_by(condition) %>%
  tidyboot_mean(diff)

post_test %>%
  ggplot(aes(x = condition, y = diff)) +
    geom_point(position = position_jitter(width = .2, height = 1), size = .05, alpha = 0.15) +
    geom_point(aes(y = empirical_stat), data = summary.df) +
    geom_errorbar(aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
                  data = summary.df, width = 0) +
  geom_hline(yintercept = 0, linetype = 'dotted') +
    theme_few() +
    scale_x_discrete(labels=c("otherGroup_first"="other -> own",
                              "ownGroup_first"="own -> other")) +
    labs(x = '', y = '# words for other vs. own (log ratio)') +
    theme(aspect.ratio = 3, axis.text.x = element_text(angle = 45,  vjust = 1, hjust=1))
```


## Look at specific cases where people provided shorter description to other group

* playerId 2iWCDAdpFvRCNSrEM gave informative descriptions to own community and then just 'those are squares and diamonds' to other group
* check possible effect of block order...

```{r}
post_test %>%
  filter(playerId %in% less_informative_to_other_group) %>%
  select(playerId, gameId, tangram, ownGroup, text) %>%
  spread(ownGroup, text) %>%
  View()
```

## Write out post-test

```{r}
post_test.raw %>%
  filter(gameId %in% full_post_test) %>%
  unite(condition, ownGroup, blockOrder, remove = F) %>%
  select(gameId, playerId, tangram, utt_length_words, ownGroup, blockOrder, condition, origText) %>%
  write_csv(here('data/comprehension/post_test_utterances.csv'))
```